---
title: "HW3"
author: "Nihaal Rahman (nkr2121)"
due date: "October 14, 2023"
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(knitr)
```

## Problem 1

Loading in the dataset. 
```{r}
library(p8105.datasets)
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row representing a single product from an instacart order. There are ID variables indicating the user, order, and product. Other 
variables of interest include order_hour_of_day and days_since_prior_order, which could be useful for seeing
trends among instacart users. In total, there are `r instacart |> select(product_id) |> distinct() |> count()` products found in `r instacart |> select(user_id, order_id) |> distinct() |> count()` orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

How many aisles are there? 
```{r}
instacart |> distinct(aisle) |> count()
```

There are `r instacart |> distinct(aisle) |> count()` aisles in this dataset. 

Which aisles are the most items ordered from?
```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```

Now making a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. 
```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |>
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Now making a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables 
                      fruits")) |> 
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

Making a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.
format this table for human readers (i.e. produce a 2 x 7 table).
```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |> 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour) |> 
  knitr::kable(digits = 2)
```

## Problem 2

Importing the data
```{r}
data("brfss_smart2010")
```

Data cleaning
```{r}
brfss_smart2010 = brfss_smart2010 |> 
  janitor::clean_names() |> 
  filter(topic == "Overall Health") |> 
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", 
                                                "Excellent"), ordered = TRUE)) 
```

In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r}
brfss_smart2010 |> 
  filter(year %in% c(2002, 2010)) |> 
  group_by(year, locationabbr) |> 
  summarize(locationdesc = n_distinct(locationdesc)) |> 
  filter(locationdesc >= 7)
```

