---
title: "HW3"
author: "Nihaal Rahman (nkr2121)"
due date: "October 14, 2023"
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
```

## Problem 1

Loading in the dataset. 
```{r}
library(p8105.datasets)
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row representing a single product from an instacart order. There are ID variables indicating the user, order, and product. Other 
variables of interest include order_hour_of_day and days_since_prior_order, which could be useful for seeing
trends among instacart users. In total, there are `r instacart |> select(product_id) |> distinct() |> count()` products found in `r instacart |> select(user_id, order_id) |> distinct() |> count()` orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

How many aisles are there? 
```{r}
instacart |> distinct(aisle) |> count()
```

There are `r instacart |> distinct(aisle) |> count()` aisles in this dataset. 

Which aisles are the most items ordered from?
```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```





