---
title: "HW3"
author: "Nihaal Rahman (nkr2121)"
due date: "October 14, 2023"
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(knitr)
library(ggrepel)
```

## Problem 1

Loading in the dataset. 
```{r}
library(p8105.datasets)
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row representing a single product from an instacart order. There are ID variables indicating the user, order, and product. Other 
variables of interest include order_hour_of_day and days_since_prior_order, which could be useful for seeing
trends among instacart users. In total, there are `r instacart |> select(product_id) |> distinct() |> count()` products found in `r instacart |> select(user_id, order_id) |> distinct() |> count()` orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

How many aisles are there? 
```{r}
instacart |> distinct(aisle) |> count()
```

There are `r instacart |> distinct(aisle) |> count()` aisles in this dataset. 

Which aisles are the most items ordered from?
```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```

Now making a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. 
```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |>
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Now making a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables 
                      fruits")) |> 
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

Making a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.
format this table for human readers (i.e. produce a 2 x 7 table).
```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |> 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour) |> 
  knitr::kable(digits = 2)
```

## Problem 2

Importing the data
```{r}
data("brfss_smart2010")
```

Data cleaning
```{r}
brfss_smart2010 = brfss_smart2010 |> 
  janitor::clean_names() |> 
  filter(topic == "Overall Health") |> 
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", 
                                                "Excellent"), ordered = TRUE)) 
```

In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r}
brfss_smart2010 |> 
  filter(year %in% c(2002, 2010)) |> 
  group_by(year, locationabbr) |> 
  summarize(locationdesc = n_distinct(locationdesc)) |> 
  filter(locationdesc >= 7)
```

Constructing a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. 
```{r}
data = brfss_smart2010 |> 
  filter(response == "Excellent") |> 
  group_by(locationabbr, year) %>%
  summarize(avg_data_value = mean(data_value, na.rm = TRUE)) |> 
  select(year, locationabbr, avg_data_value)
```

Making a “spaghetti” plot of this average value over time within a state.
```{r}
ggplot(data = data, aes(x = year, y = avg_data_value, group = locationabbr)) +
  geom_line(alpha = 0.7) 
```

Making a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r}
brfss_smart2010 |> 
  filter(locationabbr == "NY" & (year == 2010 | year == 2006)) |> 
  ggplot(aes(x = response, y = data_value, fill = response)) +
  geom_boxplot() +
  facet_grid(~year) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "Distribution of Data Value among Locations in NY State",
    x = "Response",
    y = "Data Value",
    fill = "Response"
  )
```


## Problem 3

Loading, tidying, merging, and otherwise organizing the data sets.
```{r}
covar <- read_csv("~/Downloads/nhanes_covar.csv", skip = 4)

accel <- read_csv("~/Downloads/nhanes_accel.csv")

merged_data = left_join(covar, accel, by = "SEQN") |>
  janitor::clean_names() |> 
  filter(!is.na(sex) & !is.na(age) & !is.na(bmi) & !is.na(education)) |> 
  filter(age >= 21) |> 
  mutate(sex = ifelse(sex == 1, "male", "female")) |> 
  mutate(education = ifelse(education == 1, "Less than high school", 
                            ifelse(education == 2, "High school equivalent", "More than high school"))) |> 
  mutate(education = factor(education, levels = c("Less than high school", "High school equivalent", "More than high school")))
   
```

Producing a reader-friendly table for the number of men and women in each education category
```{r}
merged_data |> 
  janitor::tabyl(sex, education)
```

Creating a visualization of the age distributions for men and women in each education category.
```{r}
merged_data |> 
  ggplot(aes(x = education, y = age, fill = education)) +
  geom_boxplot() +
  facet_grid(~sex) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(
    title = "Distribution of Age for Women and Men in Each Education Group",
    x = "Education",
    y = "Age",
    fill = "Age (yrs)"
  )
```

Using the tidied dataset, aggregating across minutes to create a total activity variable for each participant. Plotting these total activities (y-axis) against age (x-axis), while comparing men to women with separate panels for each education level, and including a smooth to illustrate differences.
```{r}
merged_data %>% 
  mutate(total_activity = rowSums(select(., starts_with("min")), na.rm = TRUE)) |> 
  ggplot(aes(x = age, y = total_activity, color = sex)) +
  facet_grid(~education) +
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE)
```

Making a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. 
```{r}
merged_data |> 
  pivot_longer(cols = starts_with("min"), names_to = "minute", values_to = "activity") |>   mutate(minute = str_remove(minute, "min")) |> 
  mutate(minute = as.numeric(minute)) |> 
  group_by(sex, minute, education) |> 
  summarize(average_activity = mean(activity, na.rm = TRUE)) |>
  ggplot(aes(x = minute, y = average_activity, color = sex)) +
  facet_grid(~education) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE)
```



